Modelling BCG vaccination in the UK

### 2019-07-12 rmds/03-getTBinR.Rmd
complete getTBinR rewrite

`getTBinR` is an R package[@RCoreTeam2019] that I developed to facilitate working with the data[@WHO:2018] collected by the World Health Organization (WHO) on the country level epidemiology of Tuberculosis (TB). All data is freely available from the WHO^[WHO TB data: https://www.who.int/tb/country/data/download/en/] and the package code is archived on Zenodo[@Abbott:2019] and Github^[GitHub: https://github.com/seabbs/getTBinR/). The aim of `getTBinR` is to allow researchers, and other interested individuals, to quickly and easily gain access to a detailed TB data set and to start using it to derive key insights. It provides a consistent set of tools that can be used to rapidly evaluate hypotheses on a widely used data set before they are explored further using more complex methods or more detailed data. The remainder of this chapter outlines the key functionality of `getTBinR 0.6.1` , and summarises the current impact.

@@

`getTBinR` provides a single user facing function for data extraction, `get_tb_burden`. This function downloads multiple data sets from the WHO, cleans variables names where required, and finally joins all data sets together. On top of the core datasets provided by default, `getTBinR` also supports importing multiple other datasets. These include data on latent TB, HIV surveillance, intervention budgets, and outcomes (see `?additional_datasets` for a full list of available datasets). To reduce unnecessary downloads, and improve performance, downloads are cached automatically. `get_tb_burden` is called by all other package functions allowing for a seamless user experience. `get_data_dict` has similar functionality to `get_tb_burden` but extracts data dictionaries rather than the underlying data. It is called by the majority of the package functions in order to provide intelligent labels.

To improve the user experience, and to facilitate intelligent labeling, `getTBinR` provides a search function for the data dictionary (`search_data_dict`). This function is able to search, using fuzzy matching, for variables, variable descriptions, key phrases, and datasets. The code below gives an example search for `country` and `e_inc_100k` (TB incidence rate) variables, along with an accompanying search for variables referencing mortality.

@@

### Plotting a comparision between country, regional and global metric values {#gettbinr-plot-sum}

@@

## Package Infrastructure

@@

In this chapter I have introduced the `getTBinR` R package. `getTBinR` facilitates downloading the most up-to-date version of multiple TB relevant data sources from the WHO, along with the accompanying data dictionaries. It also contains functions to allow easy exploration of the data via searching data dictionaries, summarising key metrics on a regional and global level, and visualising the data in a variety of highly customisable ways. In addition, it provides both a dashboard and an automated country level report that enables the global, regional, and country level picture to be quickly summarised. It was developed using R package development best practises and has been peer-reviewed.[@Abbott:2018]

As of the 1st of August 2019, `getTBinR` has been released on CRAN for over a year. It has been downloaded over 10,000 times, has a growing user base, no outstanding bug-related issues, and has received multiple updates greatly expanding the functionality available. The standalone dashboard hosted online^[Dashboard: http://www.seabbs.co.uk/shiny/ExploreGlobalTB/] has had over 3000 unique users. It has been used as a teaching aid, as an example of open science, to facilate exploratory data analysis and to provide context for other research. In this thesis it has been used extensively in Chapter \@ref(background) to provide context and was also used as a hypothesis generating tool in all other chapters. Outside of the work presented in this thesis I have used `getTBinR` extensively as a data analysis tool^[Blog: https://www.samabbott.co.uk/tags/who/], to widen the appreciation of TB as a global health problem^[Twitter:https://twitter.com/search?q=getTBinR&src=typd]^[Reddit: https://www.reddit.com/user/seabbs/posts/], and to provide contextual information for other research^[Presentation: https://www.samabbott.co.uk/what-do-we-really-know-about-bcg/presentation.html#1].

Whilst `getTBinR` is feature complete, and stable, development work continues. Future projects include: using the `shinymeta` R package to provide downloadable R code to users of the interactive application; iterating on the current automated report to improve formating and to increase the amount of information displayed. As additional WHO TB data are released they will be added to `getTBinR`.

>>>>

`getTBinR` is an R package[@RCoreTeam2019] that I developed to facilitate working with the data[@WHO:2018] collected by the World Health Organization (WHO) on the country level epidemiology of Tuberculosis (TB). All data is freely available from the WHO^[WHO TB data: https://www.who.int/tb/country/data/download/en/] and the package code is archived on Zenodo[@Abbott:2019] and Github^[GitHub: https://github.com/seabbs/getTBinR/). The aim of `getTBinR` is to allow researchers, and other interested individuals, to quickly and easily gain access to a detailed TB data set and to start using it to derive key insights. It provides a consistent set of tools that can be used to rapidly evaluate hypotheses on a widely used data set before they are explored further using more complex methods or more detailed data. Prior to the development of `getTBinR` access to the WHO data was ad-hoc and there were no standardised visualisation or summary tools. The remainder of this chapter outlines the structure, and key functionality, of `getTBinR 0.6.1` . The use of `getTBinR` in this thesis is explored as well as the use of the package in work external to this thesis and by others. Much of the work done for this chapter was code, documentation, and case study development and so is not fully captured here. The GitHub repository contains a full development history of the package, as well as providing links to the documentation and accompanying case studies.

@@

`getTBinR` provides a single user facing function for data extraction, `get_tb_burden`. This function downloads multiple data sets from the WHO, cleans variables names where required, and finally joins all data sets together. On top of the core data sets provided by default, `getTBinR` also supports importing multiple other data sets. These include data on latent TB, HIV surveillance, intervention budgets, and outcomes (see `?additional_datasets` for a full list of available data sets). To reduce unnecessary downloads, and improve performance, downloads are cached automatically. `get_tb_burden` is called by all other package functions allowing for a seamless user experience. `get_data_dict` has similar functionality to `get_tb_burden` but extracts data dictionaries rather than the underlying data. It is called by the majority of the package functions in order to provide intelligent labels.

To improve the user experience, and to facilitate intelligent labeling, `getTBinR` provides a search function for the data dictionary (`search_data_dict`). This function is able to search, using fuzzy matching, for variables, variable descriptions, key phrases, and data sets. The code below gives an example search for `country` and `e_inc_100k` (TB incidence rate) variables, along with an accompanying search for variables referencing mortality.

@@

### Plotting a comparison between country, regional and global metric values {#gettbinr-plot-sum}

@@

## Package infrastructure

@@

In this chapter I have introduced the `getTBinR` R package. `getTBinR` facilitates downloading the most up-to-date version of multiple TB relevant data sources from the WHO, along with the accompanying data dictionaries. It also contains functions to allow easy exploration of the data via searching data dictionaries, summarising key metrics on a regional and global level, and visualising the data in a variety of highly customisable ways. In addition, it provides both a dashboard and an automated country level report that enables the global, regional, and country level picture to be quickly summarised. It was developed using R package development best practices and has been peer-reviewed.[@Abbott:2018]

As of the 1st of August 2019, `getTBinR` has been released on CRAN for over a year. It has been downloaded over 10,000 times, has a growing user base, no outstanding bug-related issues, and has received multiple updates greatly expanding the functionality available. The standalone dashboard hosted online^[Dashboard: http://www.seabbs.co.uk/shiny/ExploreGlobalTB/] has had over 3000 unique users. It has been used as a teaching aid, as an example of open science, to facilitate exploratory data analysis and to provide context for other research. In this thesis it has been used extensively in Chapter \@ref(background) to provide context and was also used as a hypothesis generating tool in all other chapters. Outside of the work presented in this thesis I have used `getTBinR` extensively as a data analysis tool^[Blog: https://www.samabbott.co.uk/tags/who/], to widen the appreciation of TB as a global health problem^[Twitter:https://twitter.com/search?q=getTBinR&src=typd]^[Reddit: https://www.reddit.com/user/seabbs/posts/], and to provide contextual information for other research^[Presentation: https://www.samabbott.co.uk/what-do-we-really-know-about-bcg/presentation.html#1].

Whilst `getTBinR` is feature complete, and stable, development work continues. Future projects include: using the `shinymeta` R package to provide downloadable R code to users of the interactive application; iterating on the current automated report to improve formatting and to increase the amount of information displayed. As additional WHO TB data are released they will be added to `getTBinR`.
